---
title: "Campbell Institute Regression Workshop I"
author: "Baobao Zhang"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Univariate OLS regression

In univariate OLS regression, we are predicting a single outcome variable from a single predictor variable. 

In our example, we will use the `USJudgeRatings` dataset. This dataset contains ratings of 43 state judges on 12 dimensions.


```{r}
# Load some packages in 
library(estimatr)
library(tidyverse)
library(plotly)

# Load in the dataset that we will use
require(graphics)
head(USJudgeRatings)
```

We are going to start out with a simple linear regression model with one predictor. We will predict `RTEN` (Worthy of retention) from `CONT` (Number of contacts of lawyer with judge).

```{r}
# Fit the linear regression model
model_1_predictor <- lm(RTEN ~ INTG, data = USJudgeRatings)

# Create the plot
ggplot(USJudgeRatings, aes(x = INTG, y = RTEN)) +
  geom_point() +  # Add data points
  geom_smooth(method = "lm", se = FALSE, color = "blue") +  # Add regression line
  geom_segment(aes(xend = INTG, yend = fitted(model_1_predictor)),
               color = "red", alpha = 0.5) +  # Add residuals
  labs(title = "OLS Regression with 1 Predictor",
       x = "Intergrity rating (INTG)",
       y = "Worthy of retention (RTEN)",
       caption = "Red lines represent residuals") +
  theme_minimal()

tidy(lm_robust(RTEN ~ INTG, data = USJudgeRatings))

```

# Multiple OLS regression

In multiple OLS regression, we are predicting a single outcome variable from multiple predictor variables.

```{r}
# Fit the linear regression model
model <- lm(RTEN ~ INTG + FAMI, data = USJudgeRatings)

# Create a grid of predictor values for INTG and FAMI
INTG_seq <- seq(min(USJudgeRatings$INTG), max(USJudgeRatings$INTG), length.out = 50)
FAMI_seq <- seq(min(USJudgeRatings$FAMI), max(USJudgeRatings$FAMI), length.out = 50)
prediction_grid <- expand.grid(INTG = INTG_seq, FAMI = FAMI_seq)
prediction_grid$RTEN <- predict(model, newdata = prediction_grid)

# Reshape predictions into a matrix for plotly's surface
# Rows correspond to FAMI and columns to INTG.
z_matrix <- matrix(prediction_grid$RTEN, 
                   nrow = length(FAMI_seq), 
                   ncol = length(INTG_seq))

# Create the interactive 3D plot with the regression surface and data points
fig <- plot_ly() %>%
  add_surface(x = INTG_seq, y = FAMI_seq, z = z_matrix,
              opacity = 0.5, showscale = FALSE,
              name = 'Regression Surface') %>%
  add_markers(data = USJudgeRatings, 
              x = ~INTG, y = ~FAMI, z = ~RTEN,
              marker = list(color = 'blue', size = 3),
              name = 'Data Points')

# Compute residuals: lines from the fitted value at (INTG, FAMI) to the observed RTEN
n <- nrow(USJudgeRatings)
# Prepare vectors for line segments: for each data point, we create a segment:
# (INTG, FAMI, fitted) --> (INTG, FAMI, observed)
x_lines <- rep(NA, 3 * n)
y_lines <- rep(NA, 3 * n)
z_lines <- rep(NA, 3 * n)

# For each observation, fill the vectors with two points and an NA to break the line
x_lines[seq(1, 3*n, by = 3)] <- USJudgeRatings$INTG        # starting x (fitted)
x_lines[seq(2, 3*n, by = 3)] <- USJudgeRatings$INTG        # ending x (observed)
x_lines[seq(3, 3*n, by = 3)] <- NA                          # break

y_lines[seq(1, 3*n, by = 3)] <- USJudgeRatings$FAMI        # starting y (fitted)
y_lines[seq(2, 3*n, by = 3)] <- USJudgeRatings$FAMI        # ending y (observed)
y_lines[seq(3, 3*n, by = 3)] <- NA                          # break

z_lines[seq(1, 3*n, by = 3)] <- model$fitted.values         # starting z (fitted value)
z_lines[seq(2, 3*n, by = 3)] <- USJudgeRatings$RTEN         # ending z (observed)
z_lines[seq(3, 3*n, by = 3)] <- NA                          # break

# Add the residual lines to the plot
fig <- fig %>%
  add_trace(x = x_lines, y = y_lines, z = z_lines,
            type = 'scatter3d', mode = 'lines',
            line = list(color = 'black', width = 2),
            name = "Residuals") %>%
  layout(title = "OLS regression with 2 predictor variables",
         scene = list(xaxis = list(title = "INTG: judicial integrity"),
                      yaxis = list(title = "FAMI: familiarity with law"),
                      zaxis = list(title = "RTEN: worthy of retention")))

fig

# Run the OLS regression with robust standard errors
model_robust_se <- lm_robust(RTEN ~ INTG + FAMI, data = USJudgeRatings)

tidy(model_robust_se)

```

# Logistic regression

For this question, we will use the "2011 Canadian National Election Study, With Attitude Toward Abortion" dataset, available here:
https://vincentarelbundock.github.io/Rdatasets/csv/carData/CES11.csv

The documentation for the dataset is here: https://vincentarelbundock.github.io/Rdatasets/doc/carData/CES11.html

## Running the logistic regression

We run a logistic regression to predict whether a respondent in the survey thinks that abortion should be banned based on the following variables:

- gender: treat this variable as a binary categorical variable
- importance: the importance of religion to the participant; convert this variable to a numerical variable such that not = 0, not very = 1, somewhat = 2, very = 3
- education: the highest level of education the respondent has completed; treat this variable as a categorical variable with 6 levels
- urban: whether the participant lives in an urban area: treat this variable as a binary categorical variable

We output the results with odds ratios and 95% confidence intervals generated from robust standard errors.

```{r}
# Calculate robust standard errors using sandwich estimator
library(sandwich)
library(lmtest)
library(broom)
library(modelsummary)

# Load in the dataset
df <- read.csv("https://vincentarelbundock.github.io/Rdatasets/csv/carData/CES11.csv")

# Recoding varibles

# gender
df$gender_female_c <- ifelse(df$gender == "Female", 1, 0)
table(df$gender, df$gender_female_c)

# importance of religion 
df$importance_c <- NA
df$importance_c[df$importance == "not"] <- 0
df$importance_c[df$importance == "notvery"] <- 1
df$importance_c[df$importance == "somewhat"] <- 2
df$importance_c[df$importance == "very"] <- 3
# check that we re-coded correctly
table(df$importance, df$importance_c)

# education
# a factor with (alphabetical) levels bachelors (Bachelors degree), college (community college or technical school), higher (graduate degree), HS (high-school graduate), lessHS (less than high-school graduate), somePS (some post-secondary).

# We are ordering this factor so that the levels are in the order of education level
df$education_c <- factor(df$education,
                          levels = c("lessHS", "HS", "somePS",
                                     "college", "bachelors", "higher"))
table(df$education_c)

# urban/rural
df$urban_c <- ifelse(df$urban == "urban", 1, 0)

# abortion
df$abortion_c <- ifelse(df$abortion == "Yes", 1, 0)
table(df$abortion_c)

# Run the logistic regression
md <- glm(formula = abortion_c ~ gender_female_c + importance_c + 
            education_c + urban_c, data = df, family = "binomial")

coeftest(md, vcov = vcovHC(md, type = "HC3"))

# Get model results with confidence intervals
# Odds ratio: need to set exponentiate = TRUE
# Add in robust standard errors by vcov = vcovHC(model, type = "HC3")
md_results <- tidy(md, vcov = vcovHC(md, type = "HC3"),
                      conf.int = TRUE, conf.level = 0.95, exponentiate = TRUE)
md_results
```

## Interpreting logistic regression

Write a brief interpretation of the results above (~150 words) in terms of the odds ratios and their confidence intervals.

The logistic regression results indicate that respondents’ demographic and religious characteristics are meaningfully associated with their opinions on whether abortion should be banned. Adjusting for other variables:

- As someone's self-reported importance of random goes up 1 point on the scale, they are 3.24 times more likely to oppose abortion ($p$-value <0.001). 

- Female respondents are 0.705 times less likely to support a ban on abortion compared to male respondents ($p$-value less than 0.01). 

- In terms of education, each increased level of education (compared with the reference category of less than high school) is associated with lower odds of supporting a ban on abortion. 

- Respondents living in urban areas are 0.739 times less likely to support a ban on abortion compared to those living in rural areas ($p$-value <0.05).

## Predicted probabilities

You can use the logistic regression model to predict the probability of a respondent supporting a ban on abortion based on their demographic and religious characteristics.

```{r}

# Get the predicted probability for all the respondents in the dataset
# Make sure you set the type to "response" to get the probability of the outcome
all_respondents_predict <- predict(md, newdata = df, type = "response")

# Make a histogram of the predicted probabilities
hist(all_respondents_predict, breaks = 20,
     main = "Predicted probabilities of supporting a ban on abortion",
     xlab = "Predicted probability", ylab = "Frequency")

# Get the predicted probability for a specific respondent
# This respondent: lives in urban area, indicates that religion is not very important to them,
# has a high school education, is male

# education_c = "HS"
predict(md, newdata = data.frame(urban_c = 1,
                                 importance_c = 1,
                                 education_c = "HS",
                                 gender_female_c = 0),
        type = "response")


```
